---
title: "p8105_hw2_sw3737"
author: "Shiqi Wu"
output: github_document
date: "2023-10-04"
---

```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

```{r import}
pols_data <- read.csv("data_files_hw2/pols-month.csv")
snp_data <- read.csv("data_files_hw2/snp.csv")
unemployment_data <- read.csv("data_files_hw2/unemployment.csv")
```

```{r pols cleaning}
cleaned_pols <- pols_data %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.integer(year),
         month = month.abb[as.integer(month)],
         president = ifelse(prez_dem == 1, "dem", "gop")) %>%
  select(-day, -prez_dem, -prez_gop)
```

```{r snp cleaning}
cleaned_snp <- snp_data %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.integer(year),
         month = month.abb[as.integer(month)]) %>%
  select(year, month, everything()) %>%
  arrange(year, month)
```

```{r unemployment cleaning}
tidied_unemployment <- unemployment_data %>%
  rename(year = Year) %>%
  gather(key = "month", value = "unemployment_rate", -year) %>%
  mutate(month = match(month, month.abb))
```

```{r merging}
merged_data <- cleaned_pols %>%
  left_join(cleaned_snp, by = c("year", "month")) %>%
  left_join(tidied_unemployment, by = c("year", "month"))
```

### Data description

pols-month.csv: This dataset provides insight into the `distribution of national politicians`, whether democratic or republican, across various roles like president, governor, senator, and representative. The dataset has a temporal component indicated by the mon variable.

snp.csv: This dataset tracks the `closing values of the Standard & Poor’s stock market index`, providing a gauge on the stock market's performance over time.

unemployment.csv: This dataset focuses on `unemployment rates`, segmented by month, over multiple years. This offers a view into the country's economic health and labor market trends.

The merged dataset, merged_data, seamlessly combines data from all three sources, linking them by year and month. With columns indicating `political distribution`, `stock market performance`, and `unemployment rates`, this dataset provides a comprehensive overview, facilitating an in-depth analysis of interrelationships between politics, the stock market, and unemployment.

## Problem 2

# Clean Mr. Trash Wheel dataset

```{r Mr. Trash Wheel}
mtw_data <- read_excel("data_files_hw2/202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel",range = "A2:N586") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    year=as.numeric(year),
    homes_powered = (500 * weight_tons)/30,
    wheel_name = "Mr. Trash Wheel"
  )
```

# Clean Professor Trash Wheel Dataset
```{r Mr. Trash Wheel}
ptw_data <- read_excel("data_files_hw2/202309 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M108") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    homes_powered = (500 * weight_tons)/30,
    wheel_name = "Professor Trash Wheel"
  )
```

# Clean Gwynnda Trash Wheel Dataset
```{r Mr. Trash Wheel}
gwynnda_data <- read_excel("data_files_hw2/202309 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:L157") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    homes_powered = (500 * weight_tons)/30,
    wheel_name = "Gwynnda Trash Wheel"
  )
```

# Combine with the Mr. Trash Wheel dataset to produce a single tidy dataset

```{r combine}
trashwheel_df = bind_rows(mtw_data, ptw_data, gwynnda_data) |> select(wheel_name = "wheel_name", everything())
```

# Data description

The Trash Wheel dataset consists of observations from three unique Trash Wheels: Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. The combined dataset has a total of `r nrow(trashwheel_df)` observations and `r ncol(trashwheel_df)` variables, capturing various attributes, including `wheel_name`, `dumpster`, `month`, `year`, `date`, `weight_tons`, `volume_cubic_yard`, `plastic_bottles`, `polystyrene`, `cigarette_butts`, `glass_bottles`, `plastic_bags`, `wrappers`, `sports_balls`, and `home_powered`. A total weight of `r sum(ptw_data$weight_tons)` tons trash was collected by Professor Trash Wheel. Gwynnda collected a total of `r gwynnda_data%>%filter(month == 'July'&year == 2021)%>%summarize(sum = sum(cigarette_butts))%>%sum)` cigarette butts in July of 2021.

## Problem 3

# Importing and cleaning the baseline demographics dataset

```{r}
baseline <- read.csv("data_files_hw2/MCI_baseline.csv") %>%
  janitor::clean_names() %>%
  mutate(sex = case_match(sex,
                          1 ~ "Male",
                          0 ~ "Female"),
         sex = as.factor(sex),
         apoe4 = case_match(apoe4,
                            1 ~ "APOE4 carrier",
                            0 ~ "APOE4 non-carrier"),
         apoe4 = as.factor(apoe4)) %>%
  filter(age_at_onset == "." | current_age < age_at_onset)
```